Growth model analysis 3spp WB

### 2019-03-06 MS_Analysis_3spp_4rivers_Patterns_from 2spp.Rmd
riverOrdered and metaPop01 analysis in one chunk

# Has mean size changed over time?
means <- 
  cd %>% mutate( age = year - cohort ) %>%
    filter( ageInSamples %in% 1:10, age < 3, year < 2016, area %in% c('inside','trib') ) %>%
    group_by(species,riverOrdered,ageInSamples,year,season,age) %>%
    summarize(meanLen = mean(observedLength,na.rm = TRUE),
              sdLen = sd(observedLength,na.rm = TRUE),
              meanGR = mean(grLength,na.rm = TRUE),
              sdGR = sd(grLength,na.rm = TRUE)
              )

@@

lenSlopes <- means %>%
  group_by( species, riverOrdered, ageInSamples, season, age ) %>%
  nest()

@@

save(lenSlopesTidy,file = 'D:/projects/westBrook-patterns/data/out/lenSlopesTidy.RData')
##

@@

    geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2009,y=104, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2009,y=82, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ riverOrdered + age)

ggsave(paste0('figures/meanLengths.png'), width = 11, height = 11, units='in')

@@

# do increases (slopes) increase over age? Yes
ggplot(lenSlopesTidy %>% filter(), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point( size = 3) +
  geom_smooth(method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_wrap(~riverOrdered)

ggsave(paste0('figures/meanLengthsOverAge.png'), width = 6, height = 6, units='in')

@@

save(grSlopesTidy,file = 'D:/projects/westBrook-patterns/data/out/grSlopesTidy.RData')
##
means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

@@

    geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2011.5,y=0.66, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2011.5,y=0.56, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ riverOrdered + age)

ggsave(paste0('figures/meanGRs.png'), width = 11, height = 11, units='in')

@@

# do increases (slopes) increase over age? Yes
ggplot(grSlopesTidy %>% filter(), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point(size = 3) +
  geom_smooth(method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean growth rate over years" ) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_wrap(~riverOrdered)

ggsave(paste0('figures/meanGRsOverAge.png'), width = 6, height = 6, units='in')

>>>>

```{r mean lengths by gropuingVariable}

#####################################################
#####################################################
# SElect THIS for metaPop analysis, or
groupingVariables <- c('species','metaPop01','ageInSamples', 'year','season','age')
groupingVariablesNoYear <- c('species','metaPop01','ageInSamples','season','age')

# THIS for 4 river analysis
#groupingVariables <- c('species','riverOrdered','ageInSamples','year','season','age')
#groupingVariablesNoYear <- c('species','riverOrdered','ageInSamples','season','age')

#####################################################
#####################################################

riverGroupVar <- groupingVariables[2]

# Has mean size changed over time?
means <- 
  cd %>% mutate( age = year - cohort,
                 metaPop01 = ifelse(riverOrdered == 'wb obear', 0 ,1)) %>%
    filter( ageInSamples %in% 1:10, age < 3, year < 2016, area %in% c('inside','trib'), enc == 1
            #!(riverOrdered == 'wb mitchell' & ageInSamples > 8) #low sample sizes
            ) %>% 
    group_by_at(vars(one_of(groupingVariables))) %>%
    summarize(meanLen = mean(observedLength,na.rm = TRUE),
              sdLen = sd(observedLength,na.rm = TRUE),
              meanGR = mean(grLength,na.rm = TRUE),
              sdGR = sd(grLength,na.rm = TRUE),
              n = n()
              )

@@

lenSlopes <- means %>%
  group_by_at(vars(one_of(groupingVariablesNoYear))) %>%
  #group_by(species,metaPop01,ageInSamples,year,season,age) %>%
  nest()

@@

save(lenSlopesTidy,file = paste0('D:/projects/westBrook-patterns/data/out/lenSlopesTidy_',riverGroupVar,'.RData'))
##

@@

    geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2009,y=104, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2009,y=82, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = lenSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
 # facet_grid(seasonGG ~ riverOrdered + age)
  facet_grid(as.formula(paste("seasonGG ~", riverGroupVar, "+ age")))

ggsave(paste0('figures/meanLengths_',riverGroupVar,'.png'), width = 11, height = 11, units='in')

@@

lenSlopesTidy %>%
  group_by(species, metaPop01) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
#ggplot(lenSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8)), aes((ageInSamples),estimate, color=speciesGG)) +
ggplot(lenSlopesTidy %>% filter(ageInSamples %in% 1:9), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point( size = 3) +
  geom_smooth(method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_wrap(as.formula(paste("~", riverGroupVar)))

ggsave(paste0('figures/meanLengthsOverAge_',riverGroupVar,'.png'), width = 6, height = 6, units='in')

# just bkt
ggplot(lenSlopesTidy %>% filter(ageInSamples %in% 1:9, species == 'bkt'), aes_(~ageInSamples, ~estimate, group = as.name(riverGroupVar))) +
  geom_point(aes_(color= as.name(riverGroupVar)), size = 3) +
  geom_smooth(aes_(color= as.name(riverGroupVar)), method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication()

ggsave(paste0('figures/meanLengthsOverAgeBKT_',riverGroupVar,'.png'), width = 6, height = 6, units='in')

@@

grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

save(grSlopesTidy,file = paste0('D:/projects/westBrook-patterns/data/out/grSlopesTidy_',riverGroupVar,'.RData'))
##

##
means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

@@

    geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), aes(x=2011.5,y=0.66, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), aes(x=2011.5,y=0.56, label=paste(round(estimate,4),sig)),hjust = 0)+#, color = speciesGG)) +
#  geom_text(data = grSlopesTidy %>% filter(species == 'ats', age==2), aes(x=2012.5,y=80, label=paste(round(estimate,2),sig)),hjust = 0)+#, color = speciesGG)) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(as.formula(paste("seasonGG ~", riverGroupVar, "+ age")))

ggsave(paste0('figures/meanGRs_',riverGroupVar,'.png'), width = 11, height = 11, units='in')

@@

# do increases (slopes) increase over age? Yes
#ggplot(lenSlopesTidy %>% filter(!(riverOrdered == 'wb mitchell' & ageInSamples > 8)), aes((ageInSamples),estimate, color=speciesGG)) +
ggplot(grSlopesTidy %>% filter(ageInSamples %in% 1:9), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point( size = 3) +
  geom_smooth(method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication() +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_wrap(as.formula(paste("~", riverGroupVar)))

ggsave(paste0('figures/meanLengthsOverAge_',riverGroupVar,'.png'), width = 6, height = 6, units='in')

# just bkt
ggplot(grSlopesTidy %>% filter(ageInSamples %in% 1:9, species == 'bkt'), aes_(~ageInSamples, ~estimate, group = as.name(riverGroupVar))) +
  geom_point(aes_(color= as.name(riverGroupVar)), size = 3) +
  geom_smooth(aes_(color= as.name(riverGroupVar)), method='lm', se=FALSE) +
  scale_x_continuous(breaks = 1:10, labels = c(0,0,1,1,1,1,2,2,2,2)) +
  labs( x = "Age", y = "Slope of mean body length over years" ) +
  theme_publication()

ggsave(paste0('figures/meanLengthsOverAgeBKT_',riverGroupVar,'.png'), width = 6, height = 6, units='in')
